Assessment and implementation of resource road simulation methods for projection of cumulative effects of disturbance on wildlife

### 2023-01-19 analysis/paper/paper.Rmd
re-run metrics with above changes

```{r aspat-perf-fig, fig.height=3, fig.width=fig_widths["two"], fig.cap="Proportional difference of projected mean from observed mean road metrics using different projection method variants with a 1 ha resolution cost raster across the entire Revelstoke Timber Supply Area (“overall”) and within cutblocks (“cutblocks”). Values greater than 0 indicate overprojection of the metric and negative values indicate an underprojection. The cutblocks only method is not defined for the road density, road presence or distance to road characteristics."}
# organize data for ggplot
meanTable_long <- meanTable %>% dplyr::select(-runTime) %>% 
  pivot_longer(-c(sampleType, sampleDens, areaMean, method),
               names_to = "metric", values_to = "response")%>%
  mutate(sampleType = ifelse(sampleType == sampleDens, NA_character_, 
                             sampleType),
         sampleDens = paste(method, sampleType, sampleDens) %>% 
           factor(levels = c("mst NA centroid",
                             "NA NA observed",
                             "mst random low sample density",
                             "mst regular low sample density",
                             "mst random high sample density",
                             "mst regular high sample density",
                             "dlcp regular high sample density",
                             "NA NA klementQGIS",
                             "NA NA NA"),
                  labels = c("MST centroid", 
                             "Observed",
                             "MST random low density",
                             "MST regular low density",
                             "MST random high density",
                             "MST regular high density",
                             "DLCP regular high density",
                             "Hardy QGIS",
                             "Cutblocks only")),
         metric = factor(metric,
                         levels = c("roadDensityMean", "roadPresenceMean",
                                    "distanceToRoadMean",
                                    "forestryDisturbanceMean",
                                    "roadDisturbanceMean"),
                         labels = c("Road\ndensity", "Road\npresence",
                                    "Distance\nto road",
                                    "Forestry\ndisturbance\nfootprint",
                                    "Road\ndisturbance\nfootprint")))

@@

# get proportional difference from observed
prop_dif <- projected_values %>% 
  left_join(observed_values %>% select(areaMean, metric, response),
            by = c("areaMean", "metric"), 
            suffix = c("_proj", "_obs")) %>% 
  mutate(prop_dif = (response_proj-response_obs)/response_obs,
         areaMean = factor(areaMean, labels = c("Cutblocks", "Overall")))

@@

prop_dif %>% 
  ggplot(aes(x = sampleDens, prop_dif, fill = sampleDens))+
  geom_hline(aes(yintercept = 0), color = "grey")+
  geom_col(position = position_dodge2(preserve = "single",
                                      width = 0.75))+
  facet_grid(areaMean~ metric)+
  theme_classic()+
  scale_fill_manual(values = met.brewer(pal_nm, 8) %>% .[c(8, 1:7)],
                    name = "Sampling\nMethod", guide = "none")+
  theme(text = element_text(size = 10), axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.position = "right",
        axis.text.x=element_text(size=11, angle=50, vjust=1, hjust=1))+
  coord_flip()

@@

# performance table
perf_tbl <- matchData_sum %>% 
  pivot_wider(c(metric, sampleDens), names_from = agreement, 
              values_from = count, values_fill = 0) %>% 
  mutate(sensitivity = `Agree roaded`/(`Agree roaded` + `False negative`),
         precision = `Agree roaded`/(`Agree roaded` + `False positive`),
         F_measure = (2*precision*sensitivity)/(precision+sensitivity)) %>% 
  select(metric, sampleDens, sensitivity, precision, F_measure)

@@

perf_tbl %>% 
  pivot_longer(c(sensitivity, precision, F_measure), names_to = "perf_meas",
               values_to = "value") %>% 
  mutate(perf_meas = factor(perf_meas,
                            labels = c("F-measure", "Precision", "Sensitivity"))) %>% 
  ggplot(aes(sampleDens, value, fill = sampleDens))+
  geom_col()+
  scale_fill_manual(values = met.brewer(pal_nm, 8) %>% .[c(8, 1:7)],
                    name = "Sampling\nMethod", guide = "none")+
  facet_grid(metric ~ perf_meas)+
  coord_flip()+
  ylim(0, 1)+
  theme_classic()+
  labs(y = "Performance", x = NULL)

@@

```{r figA1, fig.height=10, fig.width=fig_widths["two"], fig.cap="Proportional difference of projected mean from observed mean road metrics using different projection methods across the entire Revelstoke Timber Supply Area (“overall”) and within cutblocks (“cutblocks”) for the fine resolutiom (1 ha) and coarse resolution (100 ha) cost rasters. Values greater than 0 indicate the projection overestimated the metric and negative values indicate an underestimation."}
# organize data for ggplot
meanTable_long <- meanTable %>% dplyr::select(-runTime) %>% 
  pivot_longer(-c(sampleType, sampleDens, areaMean, resolution, method),
               names_to = "metric", values_to = "response")%>%
  mutate(sampleType = ifelse(sampleType == sampleDens, NA_character_, 
                             sampleType),
         sampleDens = paste(method, sampleType, sampleDens) %>% 
           factor(levels = c("mst NA centroid",
                             "NA NA observed",
                             "mst random low sample density",
                             "mst regular low sample density",
                             "mst random high sample density",
                             "mst regular high sample density",
                             "dlcp regular high sample density",
                             "NA klementQGIS", 
                             "NA NA NA"),
                  labels = c("MST centroid", 
                             "Observed",
                             "MST random low density",
                             "MST regular low density",
                             "MST random high density",
                             "MST regular high density",
                             "DLCP regular high density",
                             "Hardy QGIS",
                             "Cutblocks only")),
         metric = factor(metric,
                         levels = c("roadDensityMean", "roadPresenceMean",
                                    "distanceToRoadMean",
                                    "forestryDisturbanceMean",
                                    "roadDisturbanceMean"),
                         labels = c("Road\ndensity", "Road\npresence",
                                    "Distance\nto road",
                                    "Forestry\ndisturbance\nfootprint",
                                    "Road\ndisturbance\nfootprint")))

@@

# get proportional difference from observed
prop_dif <- projected_values %>% 
  left_join(observed_values %>% select(areaMean, metric, resolution, response),
            by = c("areaMean", "metric", "resolution"), 
            suffix = c("_proj", "_obs")) %>% 
  mutate(prop_dif = (response_proj-response_obs)/response_obs,
         areaMean = factor(areaMean, labels = c("Cutblocks", "Overall")))

>>>>

```{r aspat-perf-fig, fig.height=4, fig.width=fig_widths["two"], fig.cap="Proportional difference of projected mean from observed mean road metrics using different projection method variants with a 1 ha resolution cost raster. Metrics were summarised within cutblocks, outside cutblocks, and overall. Values greater than 0 indicate overprojection of the metric and negative values indicate an underprojection. The cutblocks only method is not defined for the road density, road presence or distance to road characteristics."}
# organize data for ggplot
meanTable_long <- meanTable %>% dplyr::select(-runTime, -order) %>% 
  pivot_longer(-c(sampleType, sampleDens, areaMean, method),
               names_to = "metric", values_to = "response")%>%
  mutate(sampleType = ifelse(sampleType == sampleDens, NA_character_, 
                             sampleType),
         sampleDens = paste(method, sampleType, sampleDens) %>% 
           factor(levels = c("mst NA centroid",
                             "NA NA observed",
                             "mst random low sample density",
                             "mst regular low sample density",
                             "mst random high sample density",
                             "mst regular high sample density",
                             "dlcp regular high sample density",
                             "NA NA klementQGIS",
                             "NA NA NA"),
                  labels = c("MST centroid", 
                             "Observed",
                             "MST random low density",
                             "MST regular low density",
                             "MST random high density",
                             "MST regular high density",
                             "DLCP regular high density",
                             "Hardy QGIS",
                             "Cutblocks only")),
         metric = factor(metric,
                         levels = c("roadDensityMean", "roadPresenceMean",
                                    "distanceToRoadMean",
                                    "forestryDisturbanceMean",
                                    "roadDisturbanceMean"),
                         labels = c("Road\ndensity", "Road\npresence",
                                    "Distance\nto road",
                                    "Forestry\ndisturbance\nfootprint",
                                    "Road\ndisturbance\nfootprint")))

@@

# get proportional difference from observed
prop_dif <- projected_values %>% 
  left_join(observed_values %>% select(areaMean, metric, response),
            by = c("areaMean", "metric"), 
            suffix = c("_proj", "_obs")) %>% 
  mutate(prop_dif = (response_proj-response_obs)/response_obs,
         areaMean = factor(areaMean, labels = c("Cutblocks", "Outside\ncutblocks",
                                                "Overall")))

@@

prop_dif %>% 
  ggplot(aes(x = sampleDens, prop_dif, fill = sampleDens))+
  geom_hline(aes(yintercept = 0), color = "grey")+
  geom_col(position = position_dodge2(preserve = "single",
                                      width = 0.75))+
  facet_grid(areaMean~ metric, scales = "free")+
  theme_classic()+
  scale_fill_manual(values = met.brewer(pal_nm, 8) %>% .[c(8, 1:7)],
                    name = "Sampling\nMethod", guide = "none")+
  theme(text = element_text(size = 10), axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        legend.position = "right",
        axis.text.x=element_text(size=11, angle=50, vjust=1, hjust=1))+
  coord_flip()

@@

# performance table
perf_tbl <- matchData_sum %>% 
  pivot_wider(c(metric, sampleDens), names_from = agreement, 
              values_from = count, values_fill = 0) %>% 
  mutate(sensitivity = `Agree roaded`/(`Agree roaded` + `False negative`)
         precision = `Agree roaded`/(`Agree roaded` + `False positive`),
         F_measure = (2*precision*sensitivity)/(precision+sensitivity)) %>% 
  select(metric, sampleDens, sensitivity, precision, F_measure)

@@

perf_tbl %>% 
  pivot_longer(c(sensitivity, precision, F_measure), names_to = "perf_meas",
               values_to = "value") %>% 
  mutate(perf_meas = factor(perf_meas,
                            labels = c("F-measure", "Precision", "Sensitivity")),
         value = value  +0.0001) %>%   # adding so 0 shows in figure
  ggplot(aes(sampleDens, value, fill = sampleDens))+
  geom_col()+
  scale_fill_manual(values = met.brewer(pal_nm, 8) %>% .[c(8, 1:7)],
                    name = "Sampling\nMethod", guide = "none")+
  facet_grid(metric ~ perf_meas)+
  coord_flip()+
  ylim(0, 1.0001)+
  theme_classic()+
  labs(y = "Performance", x = NULL)

@@

```{r figA1, fig.height=12, fig.width=fig_widths["two"], fig.cap="Proportional difference of projected mean from observed mean road metrics using different projection methods for the fine resolutiom (1 ha) and coarse resolution (100 ha) cost rasters. Metrics were summarised within cutblocks, outside cutblocks, and overall. Values greater than 0 indicate the projection overestimated the metric and negative values indicate an underestimation."}
# organize data for ggplot
meanTable_long <- meanTable %>% dplyr::select(-runTime) %>% 
  pivot_longer(-c(sampleType, sampleDens, areaMean, resolution, method),
               names_to = "metric", values_to = "response")%>%
  mutate(sampleType = ifelse(sampleType == sampleDens, NA_character_, 
                             sampleType),
         sampleDens = paste(method, sampleType, sampleDens) %>% 
           factor(levels = c("mst NA centroid",
                             "NA NA observed",
                             "mst random low sample density",
                             "mst regular low sample density",
                             "mst random high sample density",
                             "mst regular high sample density",
                             "dlcp regular high sample density",
                             "NA klementQGIS", 
                             "NA NA NA"),
                  labels = c("MST centroid", 
                             "Observed",
                             "MST random low density",
                             "MST regular low density",
                             "MST random high density",
                             "MST regular high density",
                             "DLCP regular high density",
                             "Hardy QGIS",
                             "Cutblocks only")),
         metric = factor(metric,
                         levels = c("roadDensityMean", "roadPresenceMean",
                                    "distanceToRoadMean",
                                    "forestryDisturbanceMean",
                                    "roadDisturbanceMean"),
                         labels = c("Road\ndensity", "Road\npresence",
                                    "Distance\nto road",
                                    "Forestry\ndisturbance\nfootprint",
                                    "Road\ndisturbance\nfootprint")))

@@

# get proportional difference from observed
prop_dif <- projected_values %>% 
  left_join(observed_values %>% select(areaMean, metric, resolution, response),
            by = c("areaMean", "metric", "resolution"), 
            suffix = c("_proj", "_obs")) %>% 
  mutate(prop_dif = (response_proj-response_obs)/response_obs,
         areaMean = factor(areaMean, labels = c("Cutblocks", "Outside\ncutblocks",
                                                "Overall")))
